Lens Modeling in \sw

### 2013-09-09 text/tex/spaghetti.tex
more edits

In \spl the user input is an educated guess for the topography of the
arrival-time surface.

A saddle point is recognizable in a contour map because contours do
not loop around it, they tend to approach and then pull away.  There
is one contour ---the saddle-point contour--- that makes an X on the
saddle point itself.  A saddle point contour may not appear on a plot
(and Figure~\ref{fig:arriv} does not show one.  But we can tell from
the other contours where the saddle-point contour would be.

>>>>

In \spl the user begins by making an educated guess for the topography
of the arrival-time surface --- specifically, the locations of the
maxima, minima and saddle points that correspond to images of the
brightest part of the source.  Input is given by tracing a proposal
for the saddle-point contours.  Figure \ref{fig:input-spag} shows some
example inputs --- and incidentally also indicates the origin of the
name \spl.

\spl, as implemented so far, assumes that the lens is dominated by a
single galaxy, and that the centre of that galaxy is the sole maximum.
Once the maximum has been identified, we wish to characterize the rest
of the arrival-time surface in relation to it.  From geometry, any
surface that has a central maximum and goes high at the edges must
have a lima\c con-like contour, meaning a two-looped curve with one
loop inside the other.  The maximum lies within the inner loop, a
minimum lies between the two loops, and a saddle point lies at the
self-intersection of the contour.  The top panel of
Figure~\ref{fig:input-spag} shows an example: red marks a suggested
maximum, green a saddle point, and blue a minimum.  (The small pink
dots are just help sketch the proposed contour.)  The middle and lower
panels of Figure~\ref{fig:input-spag} show a more complex situation.
The minimum has split into three images: two new minima with a saddle
point in between.  In this situation the contour through the new
saddle point forms a figure of eight, its two loops enclosing the two
new minima.  The configuration of a lima\c con, possibly with a figure
of eight inside, suffices to give a preliminary description of the
arrival-time surface for nearly all cases where the lens is dominated
by one galaxy, and the source is a single small object.

### 2013-09-09 text/tex/theory.tex
more edits

>>>>

In the lensed examples in Figure~\ref{fig:arriv}, the maximum in each
case is evident, but how can we distinguish the minima from the saddle
points?  The key is to examine the contours of equal arrival time.
Contour curves loop around a minimum or a maximum, but near a saddle
point, the contours do not loop around, instead they approach and then
pull away.  The contour curve precisely level with a saddle point
forms an X at the saddle point.  In Figure~\ref{fig:arriv}, one such X
is evident.  In other cases the X contour passing through the saddle
point is not itself plotted, but we can tell the other contours where
the X would be.  These self-crossing contours will play a special role
in the following.
